---
title: "Happiness"
author: "Noah Lee"
date: "2023-06-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(sqldf)
library(corrr)
library(caret)
library(ISLR)
```

## Load the data
```{r}
whr2015 <- read.csv("2015.csv")
whr2015
```

## THE YEAR 2015 ##
Preparing the data
```{r}
#Rename the columns
colnames(whr2015)[3] = "Happiness_Rank"
colnames(whr2015)[4] = "Happiness_Score"
colnames(whr2015)[5] = "Standard_Error"
colnames(whr2015)[6] = "GDP"
colnames(whr2015)[8] = "Life_Exp"
colnames(whr2015)[10] = "Trust"
colnames(whr2015)[12] = "Dystopia_Res"
#Remove Happiness_Rank as it is not needed
whr2015 <- whr2015[,-3]
#Top and bottom 10 countries and their happiness scores
head(whr2015, n=10)
tail(whr2015, n=10)
#Where the United States stands
whr2015[whr2015$Country == 'United States',]
```
    Above are the countries with the top 10 highest and lowest happiness scores.. The top 10 is mostly made up of Western European countries, while 8 of the lowest 10 rankings nations are from Sub-Saharan Africa.
    The USA is ranked 15/158.


What is the average happiness score per region?
```{r}
region_average2015 <- sqldf("SELECT Region, AVG(Happiness_Score) FROM whr2015
                            GROUP BY Region
                            ORDER BY AVG(Happiness_Score) desc")
region_average2015
```


Graph of the relationship between gdp and happiness score
```{r}
ggplot(whr2015, aes(x = GDP, y = Happiness_Score, group = Region)) +
  geom_point(aes(col=Region)) +
  labs(y = "Happiness Score")
```
It appears that countries with a higher GDP has a higher happiness score.


Correlation matrix
```{r}
whr2015.cor <- correlate(whr2015[3:11])
whr2015.cor
```



## Create a model to prdict a nation's happiness score
Split data into train and test
```{r}
set.seed(1)
#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(whr2015), replace=T, prob=c(0.75,0.25))
train <- whr2015[sample, ]
test <- whr2015[!sample, ]
#train
#test
```


Look for statistically significant variables
```{r}
model2015 <- glm(Happiness_Score ~ Region + GDP + Family + Life_Exp + Freedom + Trust + Generosity + Dystopia_Res, data=whr2015)
summary(model2015)
```
All variables appear to be stat-sig (look at p-values and ***)



Create model for train and create a confusion matrix
```{r}
predicted <- predict(model2015, test, type="response")
confusionMatrix(data=predicted, reference = as.factor(whr2015$Happiness_Score))
```



```{r}

```











