---
title: "Happiness"
author: "Noah Lee"
date: "2023-06-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(sqldf)
library(corrr)
library(caret)
library(ISLR)
```

## World Happiness Report data for 2015
Dataset source: https://www.kaggle.com/datasets/unsdsn/world-happiness
Dataset includes information gathered for the World Happiness Report survey for the year 2015.

Load the data
```{r}
whr2015 <- read.csv("2015.csv")
```


Prepare the data
1. Rename the columns by replacing periods(.) with an underscore(_) to be able to work with sqldf functions.
2. Remove the varaible "Happiness_Rank" as the dataframe's index can be used as the same.
```{r}
#Rename the columns
colnames(whr2015)[3] = "Happiness_Rank"
colnames(whr2015)[4] = "Happiness_Score"
colnames(whr2015)[5] = "Standard_Error"
colnames(whr2015)[6] = "GDP"
colnames(whr2015)[8] = "Life_Exp"
colnames(whr2015)[10] = "Trust"
colnames(whr2015)[12] = "Dystopia_Res"
#Remove Happiness_Rank as it is not needed
whr2015 <- whr2015[,-3]

head(whr2015) #a preview of the data
```


# Overview of the data
A look into the top and bottom 10 countries, along with where the United States ranks.
```{r}
#Top and bottom 10 countries and their happiness scores
head(whr2015, n=10)
tail(whr2015, n=10)
#Where the United States stands
whr2015[whr2015$Country == 'United States',]
```
  Above are the countries with the top 10 highest and lowest happiness scores.. The top 10 is mostly made up of Western European countries, while 8 of the lowest 10 rankings nations are from Sub-Saharan Africa.
  The USA is ranked 15/158 (15th out of 158).

# Boxplot (box and whisker) of the Happiness Scores.
Individual scores per country per region are portrayed as black dots to show density/skewness.
```{r}
ggplot(data=whr2015, aes(x=Happiness_Score, y=Region, fill=Region)) +
  geom_boxplot() +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5, fill='black', binwidth = 0.2) + 
  labs(x='Happiness Score', title='Happiness Score vs. Region')

```


# What is the average happiness score per region?
```{r}
region_average2015 <- sqldf("SELECT Region, AVG(Happiness_Score) FROM whr2015
                            GROUP BY Region
                            ORDER BY AVG(Happiness_Score) desc")
region_average2015
```


# Correlation matrix
```{r}
whr2015.cor <- correlate(whr2015[3:11])
whr2015.cor
```
Based on the values of the correlation matrix, **GDP** has the strongest positive relationship with the variable **Happiness_Score**.


# Visualize the relationship from above
Graph of the relationship between variables **GDP** and **Happiness Score**.
```{r}
ggplot(whr2015, aes(x = GDP, y = Happiness_Score, group = Region)) +
  geom_point(aes(col=Region)) +
  labs(y = "Happiness Score")
```
It appears that countries with a higher GDP has a higher happiness score.





## Create a model to predict a nation's happiness score
Split data into train and test
```{r}
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(whr2015), replace=T, prob=c(0.7,0.3))
train <- whr2015[sample, ]
test <- whr2015[!sample, ]

#create the model
whr_lm <- lm(Happiness_Score ~ GDP+Family+Life_Exp+Freedom+Trust+Generosity+Dystopia_Res, data=train)
summary(whr_lm)
```

Output the predicted Happiness Scores for the test data
```{r}
predictions <- predict(whr_lm, newdata=test, type='response')
as.single(predictions)
```
Compared to the actual test dataset's Happiness Scores
```{r}
test$Happiness_Score
```
The outputs are pretty close.
